---
title: "Data Intro and EDA"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(ggplot2)
library(GGally)
library(reshape2)
library(plotly)
```
# (a) Write a half-page summary of the paper, including at least the purpose of the study, the data, the collection method, its conclusions and potential impact.

  The purpose of this study is to build an accurate cloud detection algorithm that can process Multiangle Imaging SpectroRadiometer data without human intervention. In the last 20 years, there has been a global increase in green and sustainable climate initiatives, discussions which often begin with an analysis of cloud formation. Clouds in the polar regions have the most effect on global climate, and it is difficult to discern between white snow and white clouds from satellite images. Thus, it is imperative to have reliable algorithms that accurately distinguish between snow and clouds in order to have accurate global climate models.
  The data in this study were compiled in 2002 by NASA's Terra satellite, which is equipped with a Multiangle Imaging SpectroRadiometer and collected from ten orbits over one path. The paper analyzed 57 data units containing 7,114,248 1.1-km resolution pixels, with 36 radiation measurements for each pixel. The analysts focused on the 275-m red radiation measurements to build some of the features, so the entire data set is even larger than what is expounded upon in the paper. The study includes repeated visits over time to improve the expert labeling process, which created accurate classifications to build models from.
  This study demonstrates the power of statistical models and algorithms, and their potential impact on human understanding of complex scientific problems. This research aims to help scientists better understand the flow of visible and infrared radiation in the atmosphere, subsequently studying its response on clouds and changes in the arctic climate. Accurate arctic cloud models will yield a more complete and sound global climate model. Eventually these studies will further illuminate how changing cloud properties relate to broader changes in the arctic as concentrations of carbon increase in the atmosphere. 
  

# (b) Summarize the data, i.e., % of pixels for the different classes. Plot well-labeled beautiful maps using x, y coordinates the expert labels with color of the region based on the expert labels. Do you observe some trend/pattern? Is an i.i.d. assump- tion for the samples justified for this dataset?
```{r}
#make data frames from the txt files
library(dplyr)
library(ggplot2)

images_columns <- c("y", "x","cloud_label","NDAI","SD","CORR","rad_DF","rad_CF",
                    "rad_BF","rad_AF","rad_AN" )
image1 <- read.delim("./image_data/image1.txt", sep="", col.names = images_columns)
image2 <- read.delim("./image_data/image2.txt", sep="", col.names = images_columns)
image3 <- read.delim("./image_data/image3.txt", sep="", col.names = images_columns)
```

```{r}
#calculate percentage of pixels in each class, summarize data
pixel_perc <- function(image){
  n <- nrow(image)
  cloud <- nrow(filter(image, cloud_label==1))
  nocloud <- nrow(filter(image, cloud_label==-1))
  nolabel <- nrow(filter(image, cloud_label==0))
  return(list("cloud"=round(cloud/n,2), "no cloud"=round(nocloud/n, 2), 
              "unlabelled"=round(nolabel/n,2)))
}

i1_proportions <- pixel_perc(image1)
i2_proportions <- pixel_perc(image2)
i3_proportions <- pixel_perc(image3)
pixels <- data.frame(Image1=unlist(pixel_perc(image1)), Image2=unlist(pixel_perc(image2)), Image3=unlist(pixel_perc(image3)))

pixels
```


```{r}
#images beautiful maps
colors <- c("#56B4E9", "#999999", "orange")
ggplot(image1,aes(x=x, y=y, fill=factor(cloud_label)))+
  geom_tile()+
  labs(fill= "Cloud?",
        title="Expert Labels")+
  scale_fill_manual(values=colors,
                    labels=c("No Clouds", "Unlabelled", "Clouds")) +
  theme_classic()

ggplot(image2)+
  geom_tile(aes(x=x, y=y, fill=factor(cloud_label)))+
  labs(fill= "Cloud?",
       title="Expert Labels")+
  scale_fill_manual(values=colors,
                    labels=c("No Clouds", "Unlabelled", "Clouds")) +
  theme_classic()

ggplot(image3)+
  geom_tile(aes(x=x, y=y, fill=factor(cloud_label)))+
  labs(fill= "Cloud?",
       title="Expert Labels")+
  scale_fill_manual(values=colors,
                    labels=c("No Clouds", "Unlabelled", "Clouds")) +
  theme_classic()
```


  There is a large quantity of unlabelled data, generally on the border between cloud and no-cloud areas. Clouds exist in regions much larger than one pixel, and thus are grouped spacially. 
  An iid assumption for the samples is not justifed, becuase they are not independant. The label of one pixel is dependent on that of the pixels next to it, as demonstrated by the shape of the cloud and no-cloud regions above. 

# (c) Perform a visual and quantitative EDA of the dataset, e.g., summarizing (i) pair- wise relationship between the features themselves and (ii) the relationship between the expert labels with the individual features. Do you notice differences between the two classes (cloud, no cloud) based on the radiance or other features (CORR, NDAI, SD)?

```{r}
# quantitative pairwise relationship between features
cormat1 <- round(cor(image1),3)
cormat2 <- round(cor(image2),3)
cormat3 <- round(cor(image3),3)

#I could make these just lower half if that's better
melted_cormat1 <- melt(cormat1)
ggplot(data = melted_cormat1, aes(x=Var1, y=Var2, fill=value)) + 
  geom_tile()+
  ggtitle("Variable Correlation: Image 1")+
  scale_fill_gradient2(low = "blue", high = "red", mid = "white", 
   midpoint = 0, limit = c(-1,1), space = "Lab", 
   name="Correlation Value")

melted_cormat2 <- melt(cormat2)
ggplot(data = melted_cormat2, aes(x=Var1, y=Var2, fill=value)) + 
  geom_tile()+
  ggtitle("Variable Correlation: Image 2")+
  scale_fill_gradient2(low = "blue", high = "red", mid = "white", 
   midpoint = 0, limit = c(-1,1), space = "Lab", 
   name="Correlation Value")

melted_cormat3 <- melt(cormat3)
ggplot(data = melted_cormat3, aes(x=Var1, y=Var2, fill=value)) + 
  geom_tile()+
  ggtitle("Variable Correlation: Image 3")+
  scale_fill_gradient2(low = "blue", high = "red", mid = "white", 
   midpoint = 0, limit = c(-1,1), space = "Lab", 
   name="Correlation Value")
```

  These heat maps depict the pairwise relationships between all features in each image. We notice high correlation between the radiance angles, which follows intuitively since they measure radiance in the same location from different angles. These heatmaps also illustrate the correlations between the features and cloud labels. NDAI and CORR appear to have relatively high correlation with cloud label in comparison to the other features. 

```{r}
# relationship between NDAI and cloud labels
ggplot(image1, aes(x=cloud_label, y=NDAI)) + 
  geom_violin() + 
  theme_minimal() +
  ggtitle("NDAI and Cloud Label on image1")

ggplot(image2, aes(x=cloud_label, y=NDAI)) + 
  geom_violin() + 
  theme_minimal() +
  ggtitle("NDAI and Cloud Label on image2")

ggplot(image3, aes(x=cloud_label, y=NDAI)) + 
  geom_violin() + 
  theme_minimal() +
  ggtitle("NDAI and Cloud Label on image3")
```

```{r}
# relationship between CORR and cloud labels
ggplot(image1, aes(x=cloud_label, y=CORR)) + 
  geom_violin() + 
  theme_minimal() +
  ggtitle("CORR and Cloud Label on image1")

ggplot(image2, aes(x=cloud_label, y=CORR)) + 
  geom_violin() + 
  theme_minimal() +
  ggtitle("CORR and Cloud Label on image2")

ggplot(image3, aes(x=cloud_label, y=CORR)) + 
  geom_violin() + 
  theme_minimal() +
  ggtitle("CORR and Cloud Label on image3")
```

```{r}
# relationship between SD and cloud labels
ggplot(image1, aes(x=cloud_label, y=SD)) + 
  geom_violin() + 
  theme_minimal() +
  ggtitle("SD and Cloud Label on image1")

ggplot(image2, aes(x=cloud_label, y=SD)) + 
  geom_violin() + 
  theme_minimal() +
  ggtitle("SD and Cloud Label on image2")

```

```{r}
# relationship between rad_DF and cloud labels
ggplot(image1, aes(x=cloud_label, y=rad_DF)) + 
  geom_violin() + 
  theme_minimal() +
  ggtitle("rad_DF and Cloud Label on image1")

ggplot(image2, aes(x=cloud_label, y=rad_DF)) + 
  geom_violin() + 
  theme_minimal() +
  ggtitle("rad_DF and Cloud Label on image2")

ggplot(image3, aes(x=cloud_label, y=rad_DF)) + 
  geom_violin() + 
  theme_minimal() +
  ggtitle("rad_DF and Cloud Label on image3")
```
##Do you notice differences between the two classes (cloud, no cloud) based on the radiance or other features (CORR, NDAI, SD)?
  Since the radiance angles are highly correlated with one another, their relationships to the expert labels -- and consequently their respective violin plots-- are largely similar. Thus, we only look at the violin plot of rad_DF and the expert labels, and notice some differences in the plots across images. These differences make sense because of substantially differing terrain and cloud formations between the three images, which can be seen in the images from part (b). 
  We see more differences between the violin plots for NDAI and CORR across images, but the plots for SD look the same. The SD feature is used in the paper to identify smooth surfaces in the satellite images. Given that the violin plot of SD and the expert labels hugs the x-axis with an SD value close to zero, and that we are using a very small subset of images in our project, it's likely that our images have largely smooth surfaces. CORR is used as an average linear correlation of radiation measurements at different view angles, and has noticeable differences across the three images when considered with cloud labels. NDAI is defined as the normalized difference angular index, which has the widest and largest spread when compared to cloud labels, and is consistent across the three images. 
    The correlation heatmap shows darker boxes for cloud_label and NDAI as well as cloud_label and CORR. The box for cloud_label and SD is less shaded. This follows in the respective violin plots -- the higher the correlation, the greater the area inside the violin plot.